---
title: "DysbIndex_RapidDec"
author: "Sajan Raju"
date: "2024-04-17"
output:
  pdf_document:
    toc: true
    toc_depth: '2'
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Logistic regression (Dysbiosis index with other variables)
<p> 
<b> Result summary </b>

</p>
<br>

```{r cars, echo=FALSE,  warning=FALSE,message = FALSE}
library(readxl)
library(dplyr)
library(tidyverse)
library(epiDisplay)
library(gtsummary)
```


```{r cars, echo=FALSE,  warning=FALSE,message = FALSE}
setwd("/Users/sajanruio.no/OneDrive - Universitetet i Oslo/cococmo/METAGENOMICS/Rcodes/RapidDecliners/")

metadata_new<-read_excel("~/Library/CloudStorage/OneDrive-UniversitetetiOslo/cococmo/METAGENOMICS/Rcodes/RapidDecliners/metadata_spirodata_MGI_IDs_ver2.xlsx")
# Load necessary libraries (if not already installed, install them with install.packages("package_name"))
library(stats)
#metadata_new
data <-metadata_new
 #omit NA
data <- data[!is.na(data$DysbIndex_Gevers), ]
#library(dplyr)
data$Status_smoking_fu<-as.character(data$Status_smoking_bl)
data$Ethnic<-as.character(data$Ethnic)
data$Ethnic_cat<-as.character(data$Ethnic_cat)
data <- data %>%
      mutate(Ethnicity = ifelse(Ethnic == "Caucasian","Caucasian","Others"))

```

# Rapid Decliner
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
data <- data[!is.na(data$Faster_decliner_FEV1), ]
data <- data %>%
      mutate(Faster_decliner_FEV1 = ifelse(Faster_decliner_FEV1 == "Non_faster",0,1))
```

## Crude model
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
model <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers, data = data, family="binomial")
#summary(model)
#exp(cbind("Odds ratio" = coef(model), confint.default(model, level = 0.95)))
tbl_regression(model, exponentiate = TRUE)

data$Age<- as.numeric(data$Age)

```
## Age sex bmi
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + Age + Sex + BMI, data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```
## Smoke status (2 group)
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + Status_smoking_bl , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```
## Smoke status (3 group)
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + Status_smoking_bl , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```
## IL.10
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + IL.10 , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```

## IL.1b
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + IL.1b , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```

## CD4 nadir
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
data1 <- data[!is.na(data$CD4_nadir), ]
data1$CD4_nadir<-as.numeric(data1$CD4_nadir)
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + CD4_nadir , data = data1, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```


## new covars - same as Rebekka paper
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
data <- data[!is.na(data$Status_smoking_bl), ]
data <- data[!is.na(data$Age), ]
data <- data[!is.na(data$Sex), ]
data <- data[!is.na(data$Ethnic), ]

# Age and Transm_mode_BL
model2_1 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + Age + Sex +Ethnic_cat + Status_smoking_bl , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2_1), confint.default(model2_1, level = 0.95)))
tbl_regression(model2_1, exponentiate = TRUE)
```


## HIV time and Transm mode
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
#data2 <- data[!is.na(data$HIV_time), ]
#data2 <- data2[!is.na(data$Transm_mode_BL), ]

# Age and Transm_mode_BL
model2_1 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + Age + Sex +BMI +Ethnicity + Status_smoking_bl +HIV_time + Transm_mode_BL+ IL.10 + IL.1b, data = data, family="binomial")

#exp(cbind("Odds ratio" = coef(model2_1), confint.default(model2_1, level = 0.95)))
tbl_regression(model2_1, exponentiate = TRUE)
```

## All - HIV time , Transm mode and CD4_nadir
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
#data2 <- data[!is.na(data$HIV_time), ]
#data2 <- data2[!is.na(data$Transm_mode_BL), ]
data$CD4_nadir<- as.numeric(data$CD4_nadir)
# Age and Transm_mode_BL
model2_1 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + Age + Sex +BMI +Ethnicity + Status_smoking_bl +HIV_time + Transm_mode_BL+ IL.10 + IL.1b + CD4_nadir, data = data, family="binomial")

#exp(cbind("Odds ratio" = coef(model2_1), confint.default(model2_1, level = 0.95)))
tbl_regression(model2_1, exponentiate = TRUE)
```

<br>
<br>

# Airflow limitation - Pulmonary_comorbidity  GOLD
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
data <- data[!is.na(data$Pulm_como), ]
data <- data %>%
      mutate(Pulm_como = ifelse(Pulm_como == "No",0,1))
```

## Crude model
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
model <- glm(Pulm_como ~ DysbIndex_Gevers, data = data, family="binomial")
#summary(model)
tbl_regression(model, exponentiate = TRUE)


```

## Age  
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform linear regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + Age, data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)

```
## Sex  
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform linear regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + Sex, data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)

```
## BMI  
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform linear regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + BMI, data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)

```
## Age sex bmi
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform linear regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + Age + Sex + BMI, data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)

```

## Age sex 
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform linear regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + Age + Sex , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)

```
## Age BMI 
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform linear regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + Age + BMI , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)

```
## Sex BMI 
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform linear regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + BMI + Sex , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)

```
## Ethnic
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform linear regression with age and sex as covariates
model2_1 <- glm(Pulm_como ~ DysbIndex_Gevers + Ethnic_cat  , data = data, family="binomial")

#exp(cbind("Odds ratio" = coef(model2_1), confint.default(model2_1, level = 0.95)))
tbl_regression(model2_1, exponentiate = TRUE)
```

## smoke status (2 group)
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + Current_smoker_fu , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```

## smoke status (3 group)
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + Status_smoking_fu , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```

## IL.10
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + IL.10 , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```

## IL.1b
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + IL.1b , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```

## CD4_nadir
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Pulm_como ~ DysbIndex_Gevers + CD4_nadir , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```


## HIV time and Transm mode
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
data2 <- data[!is.na(data$HIV_time), ]
data2 <- data[!is.na(data$Transm_mode_BL), ]


# Age and Transm_mode_BL
model2_1 <- glm(Pulm_como ~ DysbIndex_Gevers + HIV_time + Transm_mode_BL, data = data2, family="binomial")
#exp(cbind("Odds ratio" = coef(model2_1), confint.default(model2_1, level = 0.95)))
tbl_regression(model2_1, exponentiate = TRUE)
```


# Sensitivity analysis - ( only A+B+C=reproducible spirometry or ="good quality") 

```{r , echo=FALSE,  warning=FALSE,message = FALSE}
metadata_new<-read_excel("metadata_spirodata_MGI_IDs_ver2.xlsx")
# Load necessary libraries (if not already installed, install them with install.packages("package_name"))
library(stats)
#metadata_new
data <-metadata_new
 #omit NA
data <- data[!is.na(data$DysbIndex_Gevers), ]
#library(dplyr)
data$Status_smoking_fu<-as.character(data$Status_smoking_fu)
data$Ethnic<-as.character(data$Ethnic)
data$Ethnic_cat<-as.character(data$Ethnic_cat)

data <- data[!is.na(data$Faster_decliner_FEV1), ]
data <- data %>%
       mutate(Faster_decliner_FEV1 = ifelse(Faster_decliner_FEV1 == "Non_Rapid",0,1))

data$Age<- as.numeric(data$Age)

```

```{r , echo=FALSE,  warning=FALSE,message = FALSE}
data<-data %>% 
  filter(Quality_fu == "A"  | Quality_fu == "B"  |Quality_fu == "C"  )
#%>% 
 # select(sex, height, fastest)

data <- data[!is.na(data$Faster_decliner_FEV1), ]
#data <- data %>%
#      mutate(Faster_decliner_FEV1 = ifelse(Faster_decliner_FEV1 == "Non_Rapid",0,1))
```

## crude model
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
model <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers, data = data, family="binomial")
#summary(model)
#exp(cbind("Odds ratio" = coef(model), confint.default(model, level = 0.95)))
tbl_regression(model, exponentiate = TRUE)

```
## age sex bmi
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + Age + Sex + BMI, data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```
## smoke status (2 group)
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + Current_smoker_fu , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```
## smoke status (3 group)
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + Status_smoking_fu , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```
## IL.10
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + IL.10 , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```

## IL.1b
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# Perform log regression with age and sex as covariates
model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + IL.1b , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
tbl_regression(model2, exponentiate = TRUE)


```
## CD4_nadir
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
# data <- data[!is.na(data$CD4_nadir), ]
# data$CD4_nadir<-as.numeric(data$CD4_nadir)
# # Perform log regression with age and sex as covariates
# model2 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + CD4_nadir , data = data, family="binomial")
# #exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
# tbl_regression(model2, exponentiate = TRUE)


```

## HIV time and Transm mode
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
#data2 <- data[!is.na(data$HIV_time), ]
#data2 <- data2[!is.na(data$Transm_mode_BL), ]

# Age and Transm_mode_BL
model2_1 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers +HIV_time + Transm_mode_BL, data = data, family="binomial")

#exp(cbind("Odds ratio" = coef(model2_1), confint.default(model2_1, level = 0.95)))
tbl_regression(model2_1, exponentiate = TRUE)
```

## new covars - same as Rebekka paper
```{r , echo=FALSE,  warning=FALSE,message = FALSE}
data <- data[!is.na(data$Status_smoking_fu), ]
data <- data[!is.na(data$Age), ]
data <- data[!is.na(data$Sex), ]
data <- data[!is.na(data$Ethnic), ]

# Age and Transm_mode_BL
model2_1 <- glm(Faster_decliner_FEV1 ~ DysbIndex_Gevers + Age + Sex +Ethnic_cat + Status_smoking_fu , data = data, family="binomial")
#exp(cbind("Odds ratio" = coef(model2_1), confint.default(model2_1, level = 0.95)))
tbl_regression(model2_1, exponentiate = TRUE)
```